Geometrization of characters of tori over non-Archimedean local fields

### 2013-07-16 shanghai.tex
Switched over to tikz

% !TEX encoding = UTF-8 Unicode
\documentclass[11pt]{amsart}
\title[Geometration of characters of tori]{Geometrization of characters of tori over non-Archimedean local fields}
\usepackage[british]{babel}
\usepackage{datetime}
\date{\today}
\author{Clifton Cunningham}
\address{University of Calgary}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Pacific Institute for the Mathematical Sciences at the University of Calgary}
\email{roed.math@gmail.com}
\usepackage[utf8]{inputenc}
\renewcommand{\baselinestretch}{1.2}
\usepackage[notcite,color]{showkeys}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage[shortalphabetic]{amsrefs}
\usepackage[alphabetic]{amsrefs}
\renewcommand\MR{\relax}
\usepackage{xypic}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{yfonts}
\newcommand{\mathswab}[1]{\operatorname{\textswab{#1}}}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{manfnt}

@@

  \[
  \xymatrix{
    \bm_3^*\gcs{L} \ar[rr]^{(\bm \tight{1}{\times}{2} \id)^*\mu} \ar[d]_{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \ar[d]^{\mu \boxtimes \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \ar[rr]^{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  }
  \]
  commutes, where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;

@@

  \[
  \xymatrix{
    \Frob_{\GxxG}^* \bm^* \gcs{L} \ar[rr]^{\Frob_{\GxxG}^*\mu}
    && \Frob_{\GxxG}^*(\LxL)\\
    \ar@{=}[u] \bm^*  \Frob_{G}^* \gcs{L} \ar[d]_{\bm^* F}
    && \ar[d]^{F\boxtimes F} \ar@{=}[u] \Frob_{G}^*\gcs{L}\boxtimes \Frob_{G}^*\gcs{L}\\
    \bm^*\gcs{L} \ar[rr]^{\mu}
    && \LxL\\
  }
  \]
  commutes.
\end{enumerate}
Morphisms in $\CS(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,F)$ and
  $\cs{L'} = (\gcs{L'},\mu',F')$ are character sheaves on $G$ then
  $\Hom(\cs{L},\cs{L'})$ is the set of morphisms $\varphi : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that
  \[
  \xymatrix{
  \ar[d]_{F} \Frob_G^* \gcs{L} \ar[r]^{\Frob_G^* \varphi} & \ar[d]^{F'} \Frob_G^* \gcs{L'}
  & \text{and}& \ar[d]_{\mu} m^* \gcs{L} \ar[r]^{m^* \varphi} & \ar[d]^{\mu'} m^* \gcs{L'} \\
  \gcs{L} \ar[r]^{\varphi} & \gcs{L'}
  && \LxL \ar[r]^{\varphi\boxtimes \varphi} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}\\
  }
  \]
  both commute.
\end{enumerate}

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,F) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \CS(G) \to \CS(H)$ such that
  \[
  \xymatrix{
  \CSiso{G} \ar[r]^{f^*} \ar[d]^{t_{G}} &  \ar[d]^{t_{H}} \CSiso{H}\\
  \Hom_\text{grp}(G(\Fq),\EEx) \ar[r] & \Hom_\text{grp}(H(\Fq),\EEx)
  }
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
 \begin{equation}\label{eq:pi-2}
   \xymatrix{
     0 \ar[r] & G^0 \ar[r] & G \ar[r] & \pi_0(G) \ar[r] & 0.
    }
 \end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
 \[
  \xymatrix{
\CS(0) \ar[r] & \CS(\pi_0(G)) \ar[r] & \CS(G) \ar[r] & \CS(G^0) \ar[r] & \CS(0)
  }
 \]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
 \begin{equation}\label{eq:pi-1}
  \xymatrix{
\CSiso{0} \ar[r] & \CSiso{\pi_0(G)} \ar[r] & \CSiso{G} \ar[r] & \CSiso{G^0} \ar[r] & \CSiso{0}.
  }
 \end{equation}
It is easy to see that $\CSiso{0}$ is trivial. In this paper we will find $\CSiso{G^0}$ and $\CSiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated, and we will also see that \eqref{eq:pi-1} is split exact; this allows us to find $\CSiso{G}$. To prove this, we find it useful to offer an alternate perspective on a subcategory of $\CS(G)$, in the next section.

@@

Now we may define $\CS_b(G)$, where $G$ is a smooth group scheme locally of finite type over $\Fq$.
\begin{enumerate}
\labitem{(CSb.0)}{CSb.0}
An object in $\CS_b(G)$ is a pair $(\pi,\psi)$ where $\pi : H\to G$ is a finite etale cover in the category of group schemes over $\Fq$ such that $\Gal(\bFq/\Fq)$ acts trivially on $\ker {\bar \pi}$ (so $\ker {\bar \pi} = \ker \pi$) and $\psi : \ker \pi \to \Aut(V)$ is an irreducible representation on a $\EE$-vector space $V$.
\[
\xymatrix{
V & \Aut(V) & 1 \ar[r] & \ar@/_2pc/[ll]^{\psi} \ker \pi \ar[r] & H \ar[r]^{\pi} & G \ar[r] &1 
}
\]
\labitem{(CSb.1)}{CSb.1}
A morphism $T : (\pi,\psi) \to (\pi',\psi')$ is a linear transformation $T : V \to V'$ which, if non-trivial, has the property that there is a finite etale cover $\pi_1 : H_1 \to G$ in the category of group schemes over $\Fq$ and morphisms $H_1\to H$, $H_1\to H'$ inducing homomorphisms $\ker \pi_1 \to \ker\pi$ and $\ker\pi_1\to \ker\pi$ such that the following diagram commutes.
\[
\xymatrix{
V \ar[dd] & \Aut(V) \ar@{=}[dd] & 1 \ar[r] & \ar@/^2pc/[ll]_{\psi} \ker \pi \ar[r] & H \ar[r]^{\pi} & G \ar[r] &1\\
 &  & 1 \ar[r] & \ar[u] \ar[d] \ker \pi \ar[r] & \ar[u] \ar[d] H_1 \ar[r]^{\pi_1} & \ar@{=}[u] \ar@{=}[d] G \ar[r] &1\\
V' & \Aut(V) & 1 \ar[r] & \ar@/_2pc/[ll]^{\psi'} \ker \pi' \ar[r] & H' \ar[r]^{\pi} & G \ar[r] &1\\ 
}
\]
\end{enumerate}
We refer to $\CS_b(G)$ as the category of \cdef{bounded character sheaves on $G$}.
\Clifton{I'd love to find a way to define $\CS_b(G)$ without insisting that $\psi$ is an \emph{irreducible} representation, and then prove that the full subcategory of simple objects in $\CS_b(G)$ is a full subcategory of $\CS(G)$. I'm studying my notes from BIRS, letters from Masoud, and SGA7, to see how to do this.}
\begin{proposition}
$\CS_b(G)$ is a full subcategory of $\CS(G)$.
\end{proposition}

@@

  Let $g$ be any $\Fq$-rational point on $G$ (such as the identity for
  $G$) and let $\bg$ be the geometric point on $G$ lying above
  $g$ (we fixed $\Fq \hookrightarrow \bFq$ long ago). Since $G$ is
  geometrically connected, the geometric point $\bg$ determines
  an equivalence between the category of Weil local systems on $G$ and
  $\ell$-adic representations of $\Weil(G,\bg)$
  \cite{Deligne:Weil2}*{1.1.12}. Now let $(\gcs{L},\mu,F)$ be a
  character sheaf on $G$. Then the isomorphism class of the Weil local
  system $(\gcs{L},F)$ corresponds to an $\ell$-adic character of the
  Weil group $\Weil(G, \bg)$. The $\Fq$-rational point $g$ under
  the geometric point $\bg$ determines a splitting
  $\Weil(\bFq/\Fq)\to \Weil(G,\bg)$ of the exact sequence
  \[
  \xymatrix{
    1 \ar[r] & \pi_1(\bG, \bg) \ar[r] &  \Weil(G,\bg) \ar[r]
    & \ar@/_1pc/@{.>}[l]_{g} \Weil(\bFq/\Fq) \ar[r] & 1
    }
  \]
  In this way the Weil local system $(\gcs{L},F)$ determines an $\ell$-adic character
  \[
  t_{(\gcs{L},F)}(g) : \Weil(\bFq/\Fq) \to \EEx \qquad \text{defined by}\qquad \Frob \mapsto t^\Frob_{(\gcs{L},F)}(g)
  \]
  of the Weil group $\Weil(\bFq/\Fq)$; the image of $\Frob$ under this
  character is $t^\Frob_\cs{L}(g)$, the trace of Frobenius of $\cs{L}$ at $g$.

  On the other hand, we have already seen that $t^\Frob_\cs{L} : G(\Fq) \to \EEx$
  is a group homomorphism. Since $G$ is an algebraic group over $\Fq$, it is a
  variety over $\Fq$, so finitely generated over $\Fq$, so $G(\Fq)$ is finite;
  it follows that $t^\Frob_\cs{L}(g) = t^\Frob_{(\gcs{L},F)}(g)$ is a root of unity,
  for every $g\in G(\Fq)$.  Since $\Weil(\bFq/\Fq)$ is generated by
  $\Frob$ and $t_{(\gcs{L},F)}(g) : \Weil(\bFq/\Fq) \to \EEx$ is
  a character, it follows that the image of $t_{(\gcs{L},F)}(g)$ is a
  finite subset of $\mu(\EE)$. Thus, $t_{(\gcs{L},F)}(g)$ extends to
  an $\ell$-adic character of $\Gal(\bFq/\Fq)$, which will be denoted
  by the same symbol.
  \[
  \xymatrix{
    1 \ar[r] & \ar@{=}[d] \pi_1(\bG, \bg) \ar[r] & \ar[d] \Weil(G,\bg) \ar[r] & \ar[d]  \Weil(\bFq/\Fq) \ar[r] & 1\\
    1 \ar[r] &  \pi_1(\bG, \bg) \ar[r] & \pi_1(G,\bg) \ar[r] & \Gal(\bFq/\Fq) \ar[r] & 1\\
    }
  \]

@@

  \labitem{(cs.1)}{cs.1} an indexed set of isomorphisms
    $\mu_{x,y} \ceq (x \times y)^*\mu : \gcs{L}_{x+y} \mathop{\longrightarrow}\limits^{\iso} \gcs{L}_{x} \otimes \gcs{L}_{y}$,
    for all $x,y \in G(\bFq)$, such that
    \[
    \xymatrix{
      \gcs{L}_{x+y+z} \ar[rr]^{\mu_{x+y,z}} \ar[d]_{\mu_{x,y+z}}
      && \gcs{L}_{x+y}\otimes \gcs{L}_{z} \ar[d]^{\mu_{x,y} \otimes \id} \\
      \gcs{L}_{x} \otimes \gcs{L}_{y+z} \ar[rr]^{\id \otimes \mu_{y,z}}
      && \gcs{L}_{x} \otimes\gcs{L}_{y} \otimes \gcs{L}_{z}\\
      }
    \]
    commutes, for all $x,y,z\in G(\bFq)$; and

  \labitem{(cs.2)}{cs.2} an indexed set of isomorphisms $F_{x} \ceq x^*F: \gcs{L}_{\Frob(x)} \to \gcs{L}_x$
    \Clifton{\textdbend There's hidden subtlety here too: maybe I
      should explain which Frobenius this is and how it acts. }
    such that
    \[
    \xymatrix{
      \gcs{L}_{\Frob(x)+\Frob(y)} \ar[d]_{F_{x+y}} \ar[rr]^{\mu_{\Frob(x),\Frob(y)}}
      &&\ar[d]^{F_x\otimes F_y}  \gcs{L}_{\Frob(x)}\otimes \gcs{L}_{\Frob(y)}\\
      \gcs{L}_{x+y} \ar[rr]^{\mu_{x,y}}
      && \gcs{L}_x \otimes\gcs{L}_y\\
      }
    \]
    commutes, for all $x,y\in G(\bFq)$.
  \end{enumerate}
  Under this equivalence, a map $\varphi : \cs{L} \to \cs{L'}$ of
  character sheaves on $G$ corresponds to:
  \begin{enumerate}
  \labitem{(cs.3)}{cs.3} an indexed set $\varphi_x \ceq x^*\varphi: \gcs{L}_x \to \gcs{L'}_x$
    of linear transformations (of one-dimensional $\EE$-vector spaces),
    as $x$ runs over $G(\bFq)$, such that
    \[
    \xymatrix{
      \ar[d]_{F_x} \gcs{L}_{\Frob(x)} \ar[r]^{\varphi_{\Frob(x)}} & \ar[d]^{F_x'} \gcs{L'}_{\Frob(x)}
      &\text{and}& \ar[d]_{\mu} \gcs{L}_{x+y} \ar[r]^{\varphi_{x+y}} & \ar[d]^{\mu'_{x+y}} \gcs{L'}_{x+y} \\
      \gcs{L}_x \ar[r]^{\varphi_x} & \gcs{L'}_x
      && \gcs{L}_x\otimes\gcs{L}_y \ar[r]^{\varphi_x\otimes \varphi_y} & \gcs{L'}_x \otimes \gcs{L'}_y\\
      }
    \]
    both commute.
  \end{enumerate}
  It is clear that the category defined by conditions \ref{cs.0},
  \ref{cs.1}, \ref{cs.2} and \ref{cs.3}, denoted by
  \cdef{$\CS(G(\bFq))$}, is equivalent to $\CS(G)$. We will prove the
  proposition by exhibiting an isomorphism between
  $\CSiso{G(\bFq)}$ and $\Hh^1(W_{\Fq}, G(\bFq)^*)$.

@@

\begin{proposition}\label{prop:X}
  Let $X$ be a finitely generated abelian group equipped with an
  action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on
  $X$.  Then
  \begin{align*}
    \Hh^1(\Fq, \Hom(X, \EEx)) &\xrightarrow{\rho} \Hom(\Hh^0(\Fq, X), \EEx) \\
    \rho([z])(x) &= z(\Frob)(x)
  \end{align*}
  is an isomorphism of groups.
\end{proposition}
\begin{proof}
  We first prove that $\rho$ is surjective.  Suppose that
  $f \in \Hom(X^\Frob, \EEx).$ Since $\EEx$ is divisible
  as an abelian group it is injective as a $\ZZ$-module and thus
  $\Ext^1_{\ZZ}(X/X^\Frob, \EEx) = 0$.  Applying the functor
  $\Hom(\mbox{---}, \EEx)$ to
  $$0 \rightarrow X^\Frob \rightarrow X \rightarrow X/X^\Frob \rightarrow 0$$
  and considering the associated long exact sequence gives the
  existence of an $\tilde{f} \in \Hom(X, \EEx)$ restricting to
  $f$.  Define a $1$-cocycle $z \in \Hh^1(\Fq, \Hom(X,\EEx))$ by
  mapping $\Frob$ to $\tilde{f}$.  Then $\rho([z]) = f$.

@@

  \begin{center}
  \begin{tikzpicture}[auto]
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=3em,minimum width=2em] {
  & B \\
  I & A \\};
  \draw [right hook->] (m-2-2) -- (m-1-2);
  \draw [->] (m-2-2) to node[swap,label] {$a$} (m-2-1);
  \end{tikzpicture}
  \end{center}

  \[
  \xymatrix{
    & B \\
    I & A \ar@{>->}[u] \ar[l]_{a}
    }
  \]

  where $A$ and $B$ are arbitrary $\ZZ[\Gamma]$-modules.  Since $H_m$ acts trivially on $I$,
  $I_{H_m}A := \langle (\gamma-1)a : \gamma \in H_m, a \in A \rangle \subseteq \ker(a).$
  Since $B / I_{H_m}B \cong B \otimes_A A / I_{H_m}A$, we get

>>>>

% !TEX encoding = UTF-8 Unicode
\documentclass[11pt]{amsart}
\title[Geometration of characters of tori]{Geometrization of characters of tori over non-Archimedean local fields}
\usepackage[british]{babel}
\usepackage{datetime}
\date{\today}
\author{Clifton Cunningham}
\address{University of Calgary}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Pacific Institute for the Mathematical Sciences at the University of Calgary}
\email{roed.math@gmail.com}
\usepackage[utf8]{inputenc}
\renewcommand{\baselinestretch}{1.2}
\usepackage[notcite,color]{showkeys}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage[shortalphabetic]{amsrefs}
\usepackage[alphabetic]{amsrefs}
\renewcommand\MR{\relax}
\usepackage{xypic}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{yfonts}
\newcommand{\mathswab}[1]{\operatorname{\textswab{#1}}}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
\usepackage{manfnt}

@@

  \[
  \begin{tikzcd}[row sep=40]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  \end{tikzcd}
  \]
  commutes, where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;

@@

  \[
  \begin{tikzcd}
  \Frob_{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob_{\GxxG}^*\mu}
    && \Frob_{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob_{G}^* \gcs{L} \arrow[swap]{d}{\bm^* F}
    && \Frob_{G}^*\gcs{L}\boxtimes \Frob_{G}^*\gcs{L} \dar{F\boxtimes F} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
  commutes.
\end{enumerate}
Morphisms in $\CS(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,F)$ and
  $\cs{L'} = (\gcs{L'},\mu',F')$ are character sheaves on $G$ then
  $\Hom(\cs{L},\cs{L'})$ is the set of morphisms $\varphi : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that
  \[
  \begin{tikzcd}[column sep=40]
  \Frob_G^* \gcs{L} \rar{\Frob_G^* \varphi} \arrow[swap]{d}{F} & \Frob_G^* \gcs{L'} \dar{F'}
  & \arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \varphi} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\varphi} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\varphi\boxtimes \varphi}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
  both commute.
\end{enumerate}

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,F) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \CS(G) \to \CS(H)$ such that
  \[
  \begin{tikzcd}[row sep=40]
   \CSiso{G} \rar{f^*} \arrow[swap]{d}{t_{G}} & \CSiso{H} \dar{t_H} \\
  \Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi-2}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
\[
\begin{tikzcd}
\CS(0) \rar & \CS(\pi_0(G)) \rar & \CS(G) \rar & \CS(G^0) \rar & \CS(0)
\end{tikzcd}
\]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
\begin{equation}\label{eq:pi-1}
\begin{tikzcd}
\CSiso{0} \rar & \CSiso{\pi_0(G)} \rar & \CSiso{G} \rar & \CSiso{G^0} \rar & \CSiso{0}.
\end{tikzcd}
\end{equation}
It is easy to see that $\CSiso{0}$ is trivial. In this paper we will find $\CSiso{G^0}$ and $\CSiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated, and we will also see that \eqref{eq:pi-1} is split exact; this allows us to find $\CSiso{G}$. To prove this, we find it useful to offer an alternate perspective on a subcategory of $\CS(G)$, in the next section.

@@

Now we may define $\CS_b(G)$, where $G$ is a smooth group scheme locally of finite type over $\Fq$.
\begin{enumerate}
\labitem{(CSb.0)}{CSb.0}
An object in $\CS_b(G)$ is a pair $(\pi,\psi)$ where $\pi : H\to G$ is a finite etale cover in the category of group schemes over $\Fq$ such that $\Gal(\bFq/\Fq)$ acts trivially on $\ker {\bar \pi}$ (so $\ker {\bar \pi} = \ker \pi$) and $\psi : \ker \pi \to \Aut(V)$ is an irreducible representation on a $\EE$-vector space $V$.
\[
\begin{tikzcd}
V & \Aut(V) & 1 \rar & \arrow[bend right,swap]{ll}{\psi} \ker \pi \rar & H \rar{\pi} & G \rar &1
\end{tikzcd}
\]
\labitem{(CSb.1)}{CSb.1}
A morphism $T : (\pi,\psi) \to (\pi',\psi')$ is a linear transformation $T : V \to V'$ which, if non-trivial, has the property that there is a finite etale cover $\pi_1 : H_1 \to G$ in the category of group schemes over $\Fq$ and morphisms $H_1\to H$, $H_1\to H'$ inducing homomorphisms $\ker \pi_1 \to \ker\pi$ and $\ker\pi_1\to \ker\pi$ such that the following diagram commutes.
\[
\begin{tikzcd}
V \arrow{dd} & \Aut(V) \ar[equal]{dd} & 1 \rar & \arrow[swap,bend left]{ll}{\psi} \ker \pi \rar & H \rar{\pi} & G \rar &1\\
 &  & 1 \rar & \ker \pi_1 \rar \uar \dar & H_1 \uar \dar \rar{\pi_1} & \ar[equal]{u} \arrow[equal]{d} G \rar &1\\
V' & \Aut(V') & 1 \rar & \arrow[bend right]{ll}{\psi'} \ker \pi' \rar & H' \rar{\pi'} & G \rar &1
\end{tikzcd}
\]
\end{enumerate}
We refer to $\CS_b(G)$ as the category of \cdef{bounded character sheaves on $G$}.
\Clifton{I'd love to find a way to define $\CS_b(G)$ without insisting that $\psi$ is an \emph{irreducible} representation, and then prove that the full subcategory of simple objects in $\CS_b(G)$ is a full subcategory of $\CS(G)$. I'm studying my notes from BIRS, letters from Masoud, and SGA7, to see how to do this.}
\begin{proposition}
$\CS_b(G)$ is a full subcategory of $\CS(G)$.
\end{proposition}

@@

  Let $g$ be any $\Fq$-rational point on $G$ (such as the identity for
  $G$) and let $\bg$ be the geometric point on $G$ lying above
  $g$ (we fixed $\Fq \hookrightarrow \bFq$ long ago). Since $G$ is
  geometrically connected, the geometric point $\bg$ determines
  an equivalence between the category of Weil local systems on $G$ and
  $\ell$-adic representations of $\Weil(G,\bg)$
  \cite{Deligne:Weil2}*{1.1.12}. Now let $(\gcs{L},\mu,F)$ be a
  character sheaf on $G$. Then the isomorphism class of the Weil local
  system $(\gcs{L},F)$ corresponds to an $\ell$-adic character of the
  Weil group $\Weil(G, \bg)$. The $\Fq$-rational point $g$ under
  the geometric point $\bg$ determines a splitting
  $\Weil(\bFq/\Fq)\to \Weil(G,\bg)$ of the exact sequence
  \[
  \begin{tikzcd}
  1 \rar & \pi_1(\bG, \bg) \rar &  \Weil(G,\bg) \rar
  & \ar[swap,bend right, dotted]{l}{g} \Weil(\bFq/\Fq) \rar & 1
  \end{tikzcd}
  \]
  In this way the Weil local system $(\gcs{L},F)$ determines an $\ell$-adic character
  \[
  t_{(\gcs{L},F)}(g) : \Weil(\bFq/\Fq) \to \EEx \qquad \text{defined by}\qquad \Frob \mapsto t^\Frob_{(\gcs{L},F)}(g)
  \]
  of the Weil group $\Weil(\bFq/\Fq)$; the image of $\Frob$ under this
  character is $t^\Frob_\cs{L}(g)$, the trace of Frobenius of $\cs{L}$ at $g$.

  On the other hand, we have already seen that $t^\Frob_\cs{L} : G(\Fq) \to \EEx$
  is a group homomorphism. Since $G$ is an algebraic group over $\Fq$, it is a
  variety over $\Fq$, so finitely generated over $\Fq$, so $G(\Fq)$ is finite;
  it follows that $t^\Frob_\cs{L}(g) = t^\Frob_{(\gcs{L},F)}(g)$ is a root of unity,
  for every $g\in G(\Fq)$.  Since $\Weil(\bFq/\Fq)$ is generated by
  $\Frob$ and $t_{(\gcs{L},F)}(g) : \Weil(\bFq/\Fq) \to \EEx$ is
  a character, it follows that the image of $t_{(\gcs{L},F)}(g)$ is a
  finite subset of $\mu(\EE)$. Thus, $t_{(\gcs{L},F)}(g)$ extends to
  an $\ell$-adic character of $\Gal(\bFq/\Fq)$, which will be denoted
  by the same symbol.
  \[
  \begin{tikzcd}
  1 \rar & \ar[equal]{d} \pi_1(\bG, \bg) \rar & \Weil(G,\bg) \rar \dar & \Weil(\bFq/\Fq) \rar \dar & 1\\
  1 \rar &  \pi_1(\bG, \bg) \rar & \pi_1(G,\bg) \rar & \Gal(\bFq/\Fq) \rar & 1
  \end{tikzcd}
  \]

@@

  \labitem{(cs.1)}{cs.1} an indexed set of isomorphisms
    $\mu_{x,y} \ceq (x \times y)^*\mu : \gcs{L}_{x+y} \mathop{\longrightarrow}\limits^{\iso} \gcs{L}_{x} \otimes \gcs{L}_{y}$,
    for all $x,y \in G(\bFq)$, such that
    \[
    \begin{tikzcd}[row sep=40]
      \gcs{L}_{x+y+z} \arrow{rr}{\mu_{x+y,z}} \arrow[swap]{d}{\mu_{x,y+z}}
      && \gcs{L}_{x+y}\otimes \gcs{L}_{z} \dar{\mu_{x,y} \tight{0.5}{\otimes}{1} \id} \\
      \gcs{L}_{x} \otimes \gcs{L}_{y+z} \arrow{rr}{\id \otimes \mu_{y,z}}
      && \gcs{L}_{x} \otimes\gcs{L}_{y} \otimes \gcs{L}_{z}
    \end{tikzcd}
    \]
    commutes, for all $x,y,z\in G(\bFq)$; and

  \labitem{(cs.2)}{cs.2} an indexed set of isomorphisms $F_{x} \ceq x^*F: \gcs{L}_{\Frob(x)} \to \gcs{L}_x$
    \Clifton{\textdbend There's hidden subtlety here too: maybe I
      should explain which Frobenius this is and how it acts. }
    such that
    \[
    \begin{tikzcd}[row sep=40]
      \gcs{L}_{\Frob(x)+\Frob(y)} \arrow[swap]{d}{F_{x+y}} \arrow{rr}{\mu_{\Frob(x),\Frob(y)}}
      && \gcs{L}_{\Frob(x)}\otimes \gcs{L}_{\Frob(y)} \dar{F_x \tight{0}{\otimes}{0} F_y} \\
      \gcs{L}_{x+y} \arrow{rr}{\mu_{x,y}}
      && \gcs{L}_x \otimes\gcs{L}_y
    \end{tikzcd}
    \]
    commutes, for all $x,y\in G(\bFq)$.
  \end{enumerate}
  Under this equivalence, a map $\varphi : \cs{L} \to \cs{L'}$ of
  character sheaves on $G$ corresponds to:
  \begin{enumerate}
  \labitem{(cs.3)}{cs.3} an indexed set $\varphi_x \ceq x^*\varphi: \gcs{L}_x \to \gcs{L'}_x$
    of linear transformations (of one-dimensional $\EE$-vector spaces),
    as $x$ runs over $G(\bFq)$, such that
    \[
    \begin{tikzcd}[column sep=40]
      \arrow[swap]{d}{F_x} \gcs{L}_{\Frob(x)} \rar{\varphi_{\Frob(x)}} & \gcs{L'}_{\Frob(x)} \dar{F_x'}
      &\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}}& \arrow[swap]{d}{\mu} \gcs{L}_{x+y} \rar{\varphi_{x+y}} & \gcs{L'}_{x+y} \dar{\mu'_{x+y}} \\
      \gcs{L}_x \rar{\varphi_x} & \gcs{L'}_x
      & {} & \gcs{L}_x\otimes\gcs{L}_y \rar{\varphi_x\otimes \varphi_y} & \gcs{L'}_x \otimes \gcs{L'}_y
    \end{tikzcd}
    \]
    both commute.
  \end{enumerate}
  It is clear that the category defined by conditions \ref{cs.0},
  \ref{cs.1}, \ref{cs.2} and \ref{cs.3}, denoted by
  \cdef{$\CS(G(\bFq))$}, is equivalent to $\CS(G)$. We will prove the
  proposition by exhibiting an isomorphism between
  $\CSiso{G(\bFq)}$ and $\Hh^1(W_{\Fq}, G(\bFq)^*)$.

@@

\begin{proposition}\label{prop:X}
  Let $X$ be a finitely generated abelian group equipped with an
  action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on
  $X$.  Then
  \begin{align*}
    \Hh^1(\Fq, \Hom(X, \EEx)) &\xrightarrow{\rho} \Hom(\Hh^0(\Fq, X), \EEx) \\
    \rho([z])(x) &= z(\Frob)(x)
  \end{align*}
  is an isomorphism of groups.
\end{proposition}
\begin{proof}
  We first prove that $\rho$ is surjective.  Suppose that
  $f \in \Hom(X^\Frob, \EEx).$ Since $\EEx$ is divisible
  as an abelian group it is injective as a $\ZZ$-module and thus
  $\Ext^1_{\ZZ}(X/X^\Frob, \EEx) = 0$.  Applying the functor
  $\Hom(\mbox{---}, \EEx)$ to
  \[
  \begin{tikzcd}
  0 \rar & X^\Frob \rar & X \rar & X/X^\Frob \rar & 0
  \end{tikzcd}
  \]
  and considering the associated long exact sequence gives the
  existence of an $\tilde{f} \in \Hom(X, \EEx)$ restricting to
  $f$.  Define a $1$-cocycle $z \in \Hh^1(\Fq, \Hom(X,\EEx))$ by
  mapping $\Frob$ to $\tilde{f}$.  Then $\rho([z]) = f$.

@@

  \[
  \begin{tikzcd}
  A \rar{a} \arrow[hook]{d} & I \\
  B
  \end{tikzcd}
  \]
  where $A$ and $B$ are arbitrary $\ZZ[\Gamma]$-modules.  Since $H_m$ acts trivially on $I$,
  $I_{H_m}A := \langle (\gamma-1)a : \gamma \in H_m, a \in A \rangle \subseteq \ker(a).$
  Since $B / I_{H_m}B \cong B \otimes_A A / I_{H_m}A$, we get
  \[
  \begin{tikzcd}
  A \arrow[two heads]{r} \arrow[hook]{d} & A/I_{H_m}A \rar \arrow[hook]{d} & I \\
  B \arrow[two heads]{r} & B/I_{H_m}B \arrow[dashed]{ur} & {}
  \end{tikzcd}
  \]
  The dotted arrow exists since $I$ is injective as a
  $\ZZ[\Gamma_m]$-module, and the outer diagram shows that $I$ is
  injective as a $\ZZ[\Gamma]$-module.
